---
title: 'PSTAT 194CS Midterm Project Appendix: Code'
author: "Sunny Zhong, Isabella Sri, Bowie Chuang"
date: "2024-05-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Generating Seed Number: $R_{0}$

```{r}
# Generate 10 random numbers and take the mean of that 10 numbers to generate R_0
set.seed(111)
R_0 = round(runif(10, max = 1, min = 0), 4)
R_0 = mean(R_0)
```

### Finding $R_{1}$ and $R_{2}$: Developing the function that picks the middle four digits

```{r}
extract_middle_four = function(input_avg_number) {
  sqrd_x = input_avg_number * input_avg_number
  sqrd_x_str = as.character(sqrd_x)
  middle_start = ceiling((nchar(sqrd_x_str) - 4) / 2) + 2
  middle_digits = substr(sqrd_x_str, start = middle_start, stop = middle_start + 3)
  middle_digits = as.integer(middle_digits)
  x = format(middle_digits / 10^4, nsmall=4)
  x = as.numeric(x)
  x = round(x, digits=4)
  return(x)
}

R_1 = extract_middle_four(R_0)
R_2 = extract_middle_four(R_1)
```

### Beta Function

```{r}
beta_func = function(x, a, b) {
  beta_result = beta(a, b) * (x ^ (a - 1)) * ((1 - x) ^ (b - 1))
  return(beta_result) 
}
```

### Three Scenarios of Parameters (Using Examples from the Article):

**$\alpha > \beta; \alpha = 7, \beta = 3$**

**$\alpha = \beta; \alpha = 5, \beta = 5$**

**$\alpha < \beta; \alpha = 3, \beta = 7$**

\
*$\epsilon \in [0, 0.02]$ is constant for all three scenarios*

We will use the same seed number $R_{0}$ and the random numbers $R_{1}$ and $R_{2}$ for all three scenarios.

### Implementing RNG Method for Beta Distribution Parameters Excluding Neutrosophic Random Variables

```{r}
epsilon = c(0, 0.02)
NR_1 = R_1 - epsilon
NR_2 = R_2 - epsilon
```

$\alpha > \beta; \alpha = 7, \beta = 3$

```{r}
# a = 7, b = 3
a = 7
b = 3
M = (a - 1)/(a + b - 2)

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a, b)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```

$\alpha = \beta; \alpha = 5, \beta = 5$

```{r}
# a = 5, b = 5
a2 = 5
b2 = 5
M = (a2 - 1)/(a2 + b2 - 2)

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a2, b2)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```

$\alpha < \beta; \alpha = 3, \beta = 7$

```{r}
# a = 3, b = 7
a3 = 3
b3 = 7
M = (a3 - 1)/(a3 + b3 - 2)

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a3, b3)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```


### Implementing the RNG Method to Neutrosophic Random Variables with $\delta = 0.1$ and $\epsilon = 0.2$

```{r}
NR_1 = R_1 - epsilon
NR_2 = R_2 - epsilon

epsilon_1 = 0.2
delta = 0.1
```

$\alpha > \beta; \alpha = 7, \beta = 3$

```{r}
a_n = a + epsilon_1
b_n = b + delta
M = (a_n - 1)/(a_n + b_n - 2)
NR1_function = beta_func(NR_1, a, b)/ M

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a_n, b_n)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```


$\alpha = \beta; \alpha = 5, \beta = 5$

```{r}
a_n2 = a2 + epsilon_1
b_n2 = b2 + delta
M = (a_n2 - 1)/(a_n2 + b_n2 - 2)
NR1_function = beta_func(NR_1, a2, b2)/ M

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a_n2, b_n2)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```

$\alpha < \beta; \alpha = 3, \beta = 7$

```{r}
a_n3 = a3 + epsilon_1
b_n3 = b3 + delta
M = (a_n3 - 1)/(a_n3 + b_n3 - 2)
NR1_function = beta_func(NR_1, a2, b2)/ M

acceptance <- TRUE
for (i in 1:length(NR_2)) {
  NR1_function = beta_func(NR_1[i], a_n3, b_n3)/ M
  if (NR_2[i] > NR1_function){
    acceptance <- FALSE
    break
  }
}

if (acceptance) {
  print("We accept that N(R1) follows the beta distribution")
} else {
  print("We reject that N(R1) follows the beta distribution. Try Again with a different Random Number")
}
```

